---
title: "R Binding Conformance Checks"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(Rsassy)
```

Executable conformance checks for the R binding surface. This report compares
Rsassy outputs to the upstream `sassy` CLI where there is a matching CLI surface,
and checks R-only invariants such as list/raw inputs and threaded equivalence.
Run with:

```sh
make r-bindings-conformance
```

```{r helpers}
cli <- Sys.getenv("SASSY_CLI", unset = Sys.which("sassy"))
if (!nzchar(cli)) cli <- "/tmp/sassy-cli-root/bin/sassy"
system2(cli, "--version", stdout = TRUE)

run_cli <- function(args) {
  lines <- system2(cli, args, stdout = TRUE, stderr = TRUE)
  status <- attr(lines, "status")
  stopifnot(is.null(status) || identical(status, 0L))
  lines
}

cli_table <- function(args) {
  lines <- run_cli(args)
  lines <- lines[grepl("\t", lines)]
  utils::read.delim(text = paste(lines, collapse = "\n"), stringsAsFactors = FALSE, check.names = FALSE)
}

write_fasta <- function(records, file) {
  con <- file(file, open = "wb")
  on.exit(close(con), add = TRUE)
  for (id in names(records)) writeLines(c(paste0(">", id), records[[id]]), con, useBytes = TRUE)
  file
}

search_to_cli <- function(x, pattern_ids, text_ids) {
  data.frame(
    pat_id = pattern_ids[x$pattern_idx + 1L],
    text_id = text_ids[x$text_idx + 1L],
    cost = as.integer(x$cost),
    strand = as.character(x$strand),
    start = as.integer(x$text_start),
    end = as.integer(x$text_end),
    match_region = as.character(x$match_region),
    cigar = as.character(x$cigar),
    stringsAsFactors = FALSE
  )
}

norm_search <- function(x) {
  cols <- c("pat_id", "text_id", "cost", "strand", "start", "end", "match_region", "cigar")
  x <- as.data.frame(x)[cols]
  x$cost <- as.integer(x$cost)
  x$start <- as.integer(x$start)
  x$end <- as.integer(x$end)
  for (col in setdiff(cols, c("cost", "start", "end"))) x[[col]] <- as.character(x[[col]])
  x <- x[do.call(order, x), , drop = FALSE]
  row.names(x) <- NULL
  x
}

norm_crispr <- function(x) {
  cols <- c("guide", "text_id", "cost", "strand", "start", "end", "match_region", "cigar")
  x <- as.data.frame(x)[cols]
  x$cost <- as.integer(x$cost)
  x$start <- as.integer(x$start)
  x$end <- as.integer(x$end)
  for (col in setdiff(cols, c("cost", "start", "end"))) x[[col]] <- as.character(x[[col]])
  x <- x[do.call(order, x), , drop = FALSE]
  row.names(x) <- NULL
  x
}

tmp <- tempfile("rsassy-bindings-conf-")
dir.create(tmp)
```

## Search, encoded patterns, and SAM orientation vs CLI

```{r search-cli}
patterns <- c(p_atg = "ATG", p_ttt = "TTT")
texts <- c(t1 = "CCCCATGCCCCTTT")
pattern_fa <- write_fasta(patterns, file.path(tmp, "patterns.fa"))
text_fa <- write_fasta(texts, file.path(tmp, "texts.fa"))

cli_pairwise <- norm_search(cli_table(c("search", "--pattern-fasta", pattern_fa, "-k", "1", "--no-rc", "--alphabet", "iupac", text_fa)))
r_pairwise <- norm_search(search_to_cli(
  sassy_search(as.list(unname(patterns)), as.list(unname(texts)), 1, alphabet = "iupac", rc = FALSE, strategy = "pairwise", match_region = TRUE),
  names(patterns), names(texts)
))
cli_pairwise
r_pairwise
stopifnot(identical(r_pairwise, cli_pairwise))

cli_v2 <- norm_search(cli_table(c("search", "--pattern-fasta", pattern_fa, "-k", "1", "--no-rc", "--alphabet", "iupac", "--v2", text_fa)))
r_v2 <- norm_search(search_to_cli(
  sassy_search(as.list(unname(patterns)), as.list(unname(texts)), 1, alphabet = "iupac", rc = FALSE, strategy = "encoded_patterns", match_region = TRUE),
  names(patterns), names(texts)
))
stopifnot(identical(r_v2, cli_v2))

sam_texts <- c(chr1 = "TTTCGTTT")
sam_fa <- write_fasta(sam_texts, file.path(tmp, "sam.fa"))
cli_sam <- norm_search(cli_table(c("search", "-p", "ACGA", "-k", "0", "--alphabet", "dna", "--sam", sam_fa)))
r_sam <- norm_search(search_to_cli(
  sassy_search(list("ACGA"), as.list(unname(sam_texts)), 0, alphabet = "dna", match_region = TRUE, sam = TRUE),
  "pattern", names(sam_texts)
))
stopifnot(identical(r_sam, cli_sam))
```

## Threading and strategy equivalence

```{r threading}
texts_many <- rep("CCCC", 130L)
texts_many[c(2L, 65L, 130L)] <- c("ATGCCC", "CCATGCC", "GGGATG")
serial <- sassy_search(list("ATG"), as.list(texts_many), 0, alphabet = "dna", rc = FALSE, strategy = "pairwise", threads = 1L)
parallel <- sassy_search(list("ATG"), as.list(texts_many), 0, alphabet = "dna", rc = FALSE, strategy = "pairwise", threads = 4L)
stopifnot(identical(parallel, serial))

batch_texts <- sassy_search(list("ATG"), list("CCCCATG", "TTTATG"), 0, alphabet = "iupac", rc = FALSE, strategy = "batch_texts")
pairwise_texts <- sassy_search(list("ATG"), list("CCCCATG", "TTTATG"), 0, alphabet = "iupac", rc = FALSE, strategy = "pairwise")
stopifnot(identical(batch_texts, pairwise_texts))

batch_patterns <- sassy_search(list("ATG", "TTT"), list("CCCCATGCCCCTTT"), 0, alphabet = "iupac", rc = FALSE, strategy = "batch_patterns")
pairwise_patterns <- sassy_search(list("ATG", "TTT"), list("CCCCATGCCCCTTT"), 0, alphabet = "iupac", rc = FALSE, strategy = "pairwise")
stopifnot(identical(batch_patterns, pairwise_patterns))
```

## Raw/list input and ID mapping invariants

```{r input-invariants}
raw_hits <- sassy_search(list(charToRaw("ACGT")), list(charToRaw("TTACGTAA")), 0, alphabet = "dna", rc = FALSE)
char_hits <- sassy_search(list("ACGT"), list("TTACGTAA"), 0, alphabet = "dna", rc = FALSE)
stopifnot(identical(raw_hits, char_hits))

id_hits <- sassy_search(
  list("ATG", "TTT"),
  list("CCCCATG", "TTTATG"),
  0,
  alphabet = "iupac",
  rc = FALSE,
  pattern_id = c("guide_a", "guide_b"),
  text_id = c("record_a", "record_b")
)
stopifnot(identical(id_hits$pattern_id, c("guide_a", "guide_b", "guide_a")))
stopifnot(identical(id_hits$text_id, c("record_a", "record_b", "record_b")))
stopifnot(inherits(try(sassy_search("ATG", list("CCCCATG"), 0), silent = TRUE), "try-error"))
stopifnot(inherits(try(sassy_search(charToRaw("ATG"), list("CCCCATG"), 0), silent = TRUE), "try-error"))
```

## CRISPR vs CLI and threaded equivalence

```{r crispr-cli}
guides <- "ACGTNGG"
crispr_texts <- c(chr1 = "TTTACGTAGGTTT", chr2 = "TTTACGTAAATTT")
guide_file <- file.path(tmp, "guides.txt")
writeLines(guides, guide_file, useBytes = TRUE)
crispr_fa <- write_fasta(crispr_texts, file.path(tmp, "crispr.fa"))
crispr_out <- file.path(tmp, "crispr.tsv")
invisible(run_cli(c("crispr", "--guide", guide_file, "-k", "2", "--max-n-frac", "0.2", "--no-rc", "--threads", "1", "--output", crispr_out, crispr_fa)))
cli_crispr <- norm_crispr(utils::read.delim(crispr_out, stringsAsFactors = FALSE, check.names = FALSE))
r_crispr <- norm_crispr(sassy_crispr(as.list(guides), as.list(unname(crispr_texts)), 2, max_n_frac = 0.2, rc = FALSE, threads = 1L, text_id = names(crispr_texts)))
cli_crispr
r_crispr
stopifnot(identical(r_crispr, cli_crispr))

crispr_serial <- sassy_crispr(list("ACGTNGG", "TTTTNGG"), list("TTTACGTAGGTTT", "AAATTTTAGGAAA"), 0, rc = FALSE, threads = 1L)
crispr_parallel <- sassy_crispr(list("ACGTNGG", "TTTTNGG"), list("TTTACGTAGGTTT", "AAATTTTAGGAAA"), 0, rc = FALSE, threads = 4L)
stopifnot(identical(crispr_parallel, crispr_serial))
```

All R binding conformance checks passed.
