---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Rsassy

[![R-universe](https://sounkou-bioinfo.r-universe.dev/badges/Rsassy)](https://sounkou-bioinfo.r-universe.dev/Rsassy)

R bindings to Sassy through R's native C API. Results are returned as data
frames with 0-based coordinates and CIGAR strings.

## Install

Install from r-universe:

```r
install.packages(
  "Rsassy",
  repos = c("https://sounkou-bioinfo.r-universe.dev", "https://cloud.r-project.org")
)
```

Source installs require Cargo/rustc >= 1.91 and `xz`.

From a checkout:

```sh
R CMD INSTALL .
```

Checkout installs resolve the `sassy` Rust crate from crates.io. For CRAN/offline
source packages, run `make vendor-rust` before `R CMD build` to add
`src/rust/vendor.tar.xz`; that generated bundle is not committed by default.
Native Rsassy builds are multiversion by default: x86_64 builds install scalar,
AVX2, and AVX-512 backends; arm64 builds install scalar and NEON backends.
Rsassy selects
the best installed backend supported by the current CPU/runtime when the native
backend is first loaded. webR/WebAssembly builds use `wasm32-unknown-emscripten`
with SIMD128.

## Usage

```{r}
library(Rsassy)

sassy_search("ATCGATCG", "GGGGATCGATCGTTTT", k = 1, alphabet = "dna")
```

The result is a `sassy_matches` data frame with `pattern_idx`, `text_idx`,
`text_start`, `text_end`, `pattern_start`, `pattern_end`, `cost`, `strand`, and
`cigar`. Coordinates are 0-based and half-open.

Set `match_region = TRUE` when you also want the matched sequence. For
`strand == "-"`, `match_region` is reverse-complemented so it is in the same
direction as the input pattern and CIGAR.

```{r}
region_matches <- sassy_search(
  "ATCGATCG",
  "GGGGATCGATCGTTTT",
  k = 1,
  alphabet = "dna",
  match_region = TRUE
)
region_matches
```

The print method can color `match_region` with simple ANSI escape sequences,
following the CLI `sassy grep` alignment legend: green for matching characters,
orange for mismatches, blue for inserted text characters, and red gaps for
pattern characters absent from the text. Coloring is off by default and is meant
for ANSI-capable interactive terminals.


Reuse a searcher when making repeated calls:

```{r}
searcher <- sassy_searcher("dna")
sassy_searcher_search(searcher, "ATCGATCG", "GGGGATCGATCGTTTT", k = 1)
```

Vector inputs search every pattern against every text. Native builds can use
Rust/Rayon threads. The current Rsassy wasm32/webR build path keeps the same API
but runs the bulk loop serially until threaded webR execution is explicitly
enabled and validated.

```{r}
sassy_search(
  c("ATG", "TTT"),
  "CCCCATGCCCCTTT",
  k = 1,
  alphabet = "iupac",
  rc = FALSE,
  mode = "encoded_patterns"
)
```

`mode = "encoded_patterns"` (alias `"v2"`) is the R equivalent of CLI `--v2`
for many equal-length short patterns. `mode = "batch_patterns"` and
`mode = "encoded_patterns"` currently require `alphabet = "iupac"` and equal
pattern byte lengths.

Connections can be searched without an R-level read loop and return the same
match columns with stream-relative coordinates:

```{r}
tmp <- tempfile(fileext = ".gz")
con <- gzfile(tmp, "wb")
writeBin(charToRaw("CCCCATGCCCCTTT"), con)
close(con)

con <- gzfile(tmp, "rb")
sassy_search_connection(c("ATG", "TTT"), con, k = 1, alphabet = "iupac", rc = FALSE, mode = "encoded_patterns")
close(con)
unlink(tmp)
```

Inspect the installed build:

```{r}
sassy_features()
```

Backend loading is one-shot per R process. If you need to benchmark or debug a
specific backend, call `sassy_set_backend()` before the first native Rsassy call
in a fresh `Rscript` process. See `vignette("backend-selection", package =
"Rsassy")` for the details.

## Check

From the standalone Rsassy repository root:

```sh
make check
```

Useful targets include `make readme`, `make rd`, `make install`, `make test`,
`make vendor-rust`, and `make clean`. `make check` uses a CRAN-safe default of
two Cargo build jobs; use `make check-fast` or `make CARGO_JOBS=10 check` for
local multithreaded Cargo builds.
