---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Rsassy

[![R-universe](https://sounkou-bioinfo.r-universe.dev/badges/Rsassy)](https://sounkou-bioinfo.r-universe.dev/Rsassy)

R bindings to [Sassy](https://github.com/RagnarGrootKoerkamp/sassy) through R's
native C API. Results are returned as data frames with 0-based coordinates and
CIGAR strings.

## Install

Install from r-universe:

```r
install.packages(
  "Rsassy",
  repos = c("https://sounkou-bioinfo.r-universe.dev", "https://cloud.r-project.org")
)
```

Source installs require Cargo/rustc >= 1.91 and `xz`. Rust crates are vendored
in `src/rust/vendor.tar.xz` for offline package builds. On Linux, macOS, and
Windows, Rsassy installs multiple backend libraries when possible: scalar, AVX2,
and AVX-512 on x86_64; scalar and NEON on arm64. The webR/WebAssembly build uses
wasm SIMD128; see the [browser demo](https://sounkou-bioinfo.github.io/sassy-webR/).
Rsassy selects the best installed backend supported by the current CPU/runtime
when the backend is first loaded.

## Usage

```{r}
library(Rsassy)

sassy_search(list("ATCGATCG"), list("GGGGATCGATCGTTTT"), k = 1, alphabet = "dna")
```

The result is a `sassy_matches` data frame with `pattern_idx`, `text_idx`,
`text_start`, `text_end`, `pattern_start`, `pattern_end`, `cost`, `strand`, and
`cigar`. Coordinates are 0-based and half-open.

Set `match_region = TRUE` when you also want the matched sequence. For
`strand == "-"`, `match_region` is reverse-complemented so it is in the same
direction as the input pattern and CIGAR.

```{r}
region_matches <- sassy_search(
  list("ATCGATCG"),
  list("GGGGATCGATCGTTTT"),
  k = 1,
  alphabet = "dna",
  match_region = TRUE
)
region_matches
```

The print method can color `match_region` with simple ANSI escape sequences,
following the upstream [Sassy](https://github.com/RagnarGrootKoerkamp/sassy) CLI
`sassy grep` alignment legend: green for matching characters, orange for
mismatches, blue for inserted text characters, and red gaps for pattern
characters absent from the text. Coloring is off by default and is meant for
ANSI-capable interactive terminals.


Reuse a searcher when making repeated calls:

```{r}
searcher <- sassy_searcher("dna")
sassy_searcher_search(searcher, list("ATCGATCG"), list("GGGGATCGATCGTTTT"), k = 1)
```

List inputs search every pattern against every text. Each element can be a raw
vector or a character scalar, which also leaves room for ALTREP-backed batches
as lists. For larger batches, use `threads > 1`.

```{r}
sassy_search(
  list("ATG", "TTT"),
  list("CCCCATGCCCCTTT"),
  k = 1,
  alphabet = "iupac",
  rc = FALSE,
  strategy = "encoded_patterns"
)
```

`strategy = "encoded_patterns"` (alias `"v2"`) is the R equivalent of CLI `--v2`
for many equal-length short patterns. `batch_patterns` and `encoded_patterns`
use [Sassy](https://github.com/RagnarGrootKoerkamp/sassy)'s multi-pattern
encoding, which in `sassy` 0.2.1 is implemented for IUPAC and equal byte-length
patterns. The default `strategy = "pairwise"` is the general path for other
alphabets and mixed pattern lengths.

CLI-compatible orientation is available with `sam = TRUE`. This formats
reverse-strand `match_region` and `cigar` in text direction, matching upstream
[Sassy](https://github.com/RagnarGrootKoerkamp/sassy) `sassy --sam` output.

```{r}
sassy_search(
  list("ACGA"),
  list("TTTCGTTT"),
  k = 0,
  alphabet = "dna",
  match_region = TRUE,
  sam = TRUE
)
```

Chunked FASTA/FASTQ iteration is available with `sassy_fastx_iter()` and
`sassy_fastx_next()`. Batches expose record IDs as an ALTREP character vector and
sequences as a list of raw ALTREP slices over immutable native buffers, so search
can consume file records without first materializing sequence strings in R.

```{r}
fq <- tempfile(fileext = ".fastq")
writeLines(c("@r1", "ACGT", "+", "!!!!"), fq, useBytes = TRUE)
it <- sassy_fastx_iter(fq, batch_records = 1)
batch <- sassy_fastx_next(it)
sassy_search(list("ACG"), batch$seq, k = 0, alphabet = "dna", rc = FALSE, text_id = batch$id)
```

CRISPR guide search is available for in-memory sequences with `sassy_crispr()`.
Guides include the PAM suffix; by default the PAM must match exactly under IUPAC
matching.

```{r}
sassy_crispr(list("ACGTNGG"), list("TTTACGTAGGTTT"), k = 0, rc = FALSE)
```

For file-oriented colored `grep`, FASTA/FASTQ filtering, and large command-line
pipelines, use the upstream [Sassy](https://github.com/RagnarGrootKoerkamp/sassy)
CLI directly.

Inspect the installed build:

```{r}
sassy_features()
```

Backend loading is one-shot per R process. If you need to benchmark or debug a
specific backend, call `sassy_set_backend()` before the first native Rsassy call
in a fresh `Rscript` process. See `vignette("backend-selection", package =
"Rsassy")` for the details.

## Development

Common development commands from the repository root:

```sh
make vendor-rust  # refresh src/rust/vendor.tar.xz after Rust dependency changes
make rd           # regenerate NAMESPACE and man/*.Rd from R/search.R
make readme       # regenerate README.md from README.Rmd
make install      # install the package locally
make test         # run tinytest tests
make check        # build and run R CMD check
make reports      # render committed conformance/performance markdown reports
make clean        # remove generated build artifacts
```

`make check` uses a CRAN-safe default of two Cargo build jobs. Use
`make check-fast` or `make CARGO_JOBS=10 check` for local multithreaded Cargo
builds.
