---
title: "API and Coordinates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{API and Coordinates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(Rsassy)
```

## Functions

- `sassy_search()` searches once.
- `sassy_searcher()` creates a reusable searcher.
- `sassy_searcher_search()` searches with that object.
- `sassy_features()` prints backend/build information.
- `sassy_set_backend()` selects a backend before first use.

## Coordinates

Coordinates are 0-based and half-open.

- `text_start`, `text_end`: interval in the text.
- `pattern_start`, `pattern_end`: interval in the pattern.
- `pattern_idx`, `text_idx`: input index for vector/list inputs.

```{r}
matches <- sassy_search(list("ACGT"), list("TTACGTAA"), k = 0, alphabet = "dna", rc = FALSE)
matches[, c("text_start", "text_end", "pattern_start", "pattern_end", "cigar")]
```

## Inputs

`pattern` and `text` are lists of sequence elements. Each element may be a raw
vector or a non-missing character scalar. This keeps one raw vector as one
sequence of bytes and lets callers mix raw bytes, ordinary strings, and
ALTREP-backed string elements in the same input list.

```{r}
sassy_search(list(charToRaw("ACGT")), list(charToRaw("TTACGTAA")), k = 0, alphabet = "dna")
```

## Match regions

`match_region = TRUE` adds the matched text interval. Reverse-strand regions are
reverse-complemented to match the pattern direction.

```{r}
sassy_search(
  list("ATCGATCG"),
  list("GGGGATCGATCGTTTT"),
  k = 1,
  alphabet = "dna",
  match_region = TRUE
)
```

## Search strategies

The default `strategy = "pairwise"` searches each pattern/text pair independently.
This is the general path and works with mixed pattern lengths and all alphabets.
Other strategies are performance-oriented paths that call different Sassy kernels:

- `batch_texts`: one pattern, multiple texts per batch.
- `batch_patterns`: multiple equal-length patterns per batch.
- `encoded_patterns` / `v2`: Sassy encoded-pattern path.

`batch_patterns` and `encoded_patterns` use Sassy's multi-pattern encoding. In
`sassy` 0.2.1 that encoding is implemented for the IUPAC profile and equal
byte-length patterns.
