---
title: "API and Coordinates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{API and Coordinates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

`Rsassy` keeps the R surface small and close to the Rust API.

## Main functions

- `sassy_search()` is the convenience wrapper for one-off calls.
- `sassy_searcher()` creates a reusable searcher.
- `sassy_searcher_search()` searches with a reusable searcher.
- `sassy_search_connection()` streams from an R connection.
- `sassy_features()` reports selected backend and CPU/build features.
- `sassy_set_backend()` requests a backend before the first native call.

## Coordinates

Match coordinates are 0-based and half-open:

- `text_start`, `text_end` describe the interval in the input text.
- `pattern_start`, `pattern_end` describe the interval in the input pattern.
- `pattern_idx`, `text_idx` are 0-based indices into vector/list inputs.

This mirrors the Rust crate and avoids repeated conversion at the native
boundary.

```{r}
matches <- sassy_search("ACGT", "TTACGTAA", k = 0, alphabet = "dna", rc = FALSE)
matches[, c("text_start", "text_end", "pattern_start", "pattern_end", "cigar")]
```

## Input types

`pattern` and `text` can be character vectors, raw vectors, or lists of raw
vectors / character scalars. Use raw vectors when input is already byte-oriented
and should not pass through R string encoding.

```{r}
sassy_search(charToRaw("ACGT"), charToRaw("TTACGTAA"), k = 0, alphabet = "dna")
```

## Match regions

Set `match_region = TRUE` to include the matched text interval. Reverse-strand
regions are reverse-complemented so the region and CIGAR are in the input
pattern direction.

```{r}
sassy_search(
  "ATCGATCG",
  "GGGGATCGATCGTTTT",
  k = 1,
  alphabet = "dna",
  match_region = TRUE
)
```

## Search modes

`mode = "single"` searches each pattern/text pair independently.
`mode = "batch_texts"`, `mode = "batch_patterns"`, and
`mode = "encoded_patterns"` expose Sassy's batched paths where applicable.
The batched pattern modes currently require equal pattern byte lengths and the
IUPAC profile.
