---
title: "API and Coordinates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{API and Coordinates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(Rsassy)
```

## Functions

- `sassy_search()` searches once.
- `sassy_searcher()` creates a reusable searcher.
- `sassy_searcher_search()` searches with that object.
- `sassy_search_connection()` reads from an R connection.
- `sassy_features()` prints backend/build information.
- `sassy_set_backend()` selects a backend before first use.

## Coordinates

Coordinates are 0-based and half-open.

- `text_start`, `text_end`: interval in the text.
- `pattern_start`, `pattern_end`: interval in the pattern.
- `pattern_idx`, `text_idx`: input index for vector/list inputs.

```{r}
matches <- sassy_search("ACGT", "TTACGTAA", k = 0, alphabet = "dna", rc = FALSE)
matches[, c("text_start", "text_end", "pattern_start", "pattern_end", "cigar")]
```

## Inputs

`pattern` and `text` may be character vectors, raw vectors, or lists of raw
vectors / character scalars. Use raw vectors for byte data.

```{r}
sassy_search(charToRaw("ACGT"), charToRaw("TTACGTAA"), k = 0, alphabet = "dna")
```

## Match regions

`match_region = TRUE` adds the matched text interval. Reverse-strand regions are
reverse-complemented to match the pattern direction.

```{r}
sassy_search(
  "ATCGATCG",
  "GGGGATCGATCGTTTT",
  k = 1,
  alphabet = "dna",
  match_region = TRUE
)
```

## Modes

- `single`: independent pairwise searches.
- `batch_texts`: one pattern, multiple texts per batch.
- `batch_patterns`: multiple equal-length patterns per batch.
- `encoded_patterns` / `v2`: Sassy encoded-pattern path.

`batch_patterns` and `encoded_patterns` use Sassy's multi-pattern encoding. In
`sassy` 0.2.1 that encoding is implemented for the IUPAC profile and equal
byte-length patterns. Use `single` for other alphabets or mixed pattern lengths.
