---
title: "Backend Selection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Backend Selection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

Native `Rsassy` installs a small R/C shim plus Rust backend libraries. The C
shim selects one backend per R process.

## Default selection

With the default `"auto"` setting, Rsassy chooses the best installed backend
supported by the current runtime:

1. WebAssembly SIMD128 on wasm builds
2. AVX-512 on supported x86_64 machines
3. AVX2 on supported x86_64 machines
4. NEON on supported arm64 machines
5. scalar fallback

Inspect the installed package with:

```{r}
features <- sassy_features()
features[c(
  "rsassy_dispatch",
  "rsassy_selected_backend",
  "rsassy_installed_backends",
  "rsassy_supported_backends"
)]
```

## Explicit backend requests

Use `sassy_set_backend()` before the first native Rsassy call when benchmarking
or debugging a specific backend.

```{r}
library(Rsassy)
sassy_set_backend("scalar")
sassy_features()$rsassy_selected_backend
```

Backend loading is intentionally one-shot. `Rsassy` does not unload and reload
native backend DLLs in-process because doing so is not portable across R
platforms. Benchmark different backends in separate fresh `Rscript` processes.

## Why not one static library?

The upstream `sassy` crate selects SIMD behavior through crate/build-level
configuration and SIMD-dependent types. That makes a single function-pointer
multiversioning layer harder than in small demos where every backend is an
explicit leaf function. The current package therefore uses separate backend
libraries for native platforms and a static Rust path for webR/wasm.
