---
title: "Searching R Connections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Searching R Connections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

`sassy_search_connection()` searches an already-open readable R connection. This
avoids an R-level `readBin()` loop and keeps a native overlap window so matches
can cross chunk boundaries.

## Basic workflow

```{r}
tmp <- tempfile(fileext = ".gz")
con <- gzfile(tmp, "wb")
writeBin(charToRaw("CCCCATGCCCCTTT"), con)
close(con)

con <- gzfile(tmp, "rb")
sassy_search_connection(
  c("ATG", "TTT"),
  con,
  k = 1,
  alphabet = "iupac",
  rc = FALSE,
  mode = "encoded_patterns"
)
close(con)
unlink(tmp)
```

The returned coordinates are relative to the full stream, not to individual
chunks.

## Chunking and overlap

`chunk_size` controls how many new bytes are read per native chunk. If
`overlap = NULL`, the C shim computes a conservative overlap from the longest
pattern, `k`, and the overhang setting.

```{r}
con <- rawConnection(charToRaw("AAAAACGTAAAA"), "rb")
sassy_search_connection(
  "ACGT",
  con,
  k = 0,
  alphabet = "dna",
  rc = FALSE,
  chunk_size = 6
)
close(con)
```

Open compressed connections, such as `gzfile()`, in binary mode when possible.
