---
title: "Getting Started with Rsassy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with Rsassy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`Rsassy` provides R bindings to the Rust `sassy` approximate string matcher.
It searches short patterns in DNA, IUPAC, or ASCII text.

## Install

```{r, eval=FALSE}
install.packages(
  "Rsassy",
  repos = c("https://sounkou-bioinfo.r-universe.dev", "https://cloud.r-project.org")
)
```

## Search

```{r}
library(Rsassy)

sassy_search(list("ATCGATCG"), list("GGGGATCGATCGTTTT"), k = 1, alphabet = "dna")
```

The result is a `sassy_matches` data frame. Coordinates are 0-based and
half-open.

## Reuse a searcher

```{r}
searcher <- sassy_searcher("dna", rc = TRUE)
sassy_searcher_search(searcher, list("ATCGATCG"), list("GGGGATCGATCGTTTT"), k = 1)
```

## Multiple patterns or texts

List inputs search every pattern against every text. Each list element may be a
raw vector or a non-missing character scalar, so callers can mix byte strings,
regular strings, and ALTREP-backed string elements. `pattern_idx` and `text_idx`
identify the input indices.

```{r}
sassy_search(
  list("ATG", charToRaw("TTT")),
  list("CCCCATGCCCCTTT"),
  k = 1,
  alphabet = "iupac",
  rc = FALSE,
  strategy = "encoded_patterns"
)
```
